Replicator-mutator and noise-perturbed replicator dynamics

### 2014-09-18 paper/03_simulations.tex
iterated diffusion; plots

>>>>

\begin{itemize}
  \item which confusion matrices: normalized Nosofsky (positive matrices)
  \item repeated applications
  \item diffusion leads to a smoothing out of sender and receiver
    behavior to neighboring states
  \item how much depends on the level of confusion (see Figure~\ref{fig:diffusion-SenRec})
  \item in the limit, this will lead sender and receiver to do the
    same thing in each choice point (states for the sender, messages
    for the receiver); this follows from the XYZ theorem
\end{itemize}

\begin{figure}
  \centering

  \includegraphics[width=\textwidth]{plots/confusion_sender.pdf}

  \includegraphics[width=\textwidth]{plots/confusion_receiver.pdf}

  \caption{Iterated diffusion for arbitraty sender (top) and receiver
    strategy (bottom), with 50 states, 2 messages. Each row
    corresponds to a different confusion level $\alpha$, each colum to
    a different number of applications of iterated diffusion.}
  \label{fig:confusion-SenRec}
\end{figure}

### 2014-09-18 snippets/iteration_diffusion.Rmd
iterated diffusion; plots

```{r}
alpha= 0.1
rounds = 10000
ns = 60
nm = 2
states = seq(from=0,to=1,length.out = ns)

@@

Simple Diffusion
----------------

Let's plot an arbitrary sender and receiver strategy, and apply some noise.

First a function to plot an arbitrary sender and receiver strategy

@@

plotSender = function(sen){
  senDF = melt(sen,varnames = c("state.index","message"), value.name = "probability")
  senDF$states = states[senDF[,1]]
  senDF$message = factor(senDF$message)
  ggplot(senDF, aes(x = states, y=probability, color=message)) + geom_point() + geom_line()
}

>>>>

```{r}
alpha= 0.1
rounds = 10000
ns = 60
nm = 2

get.states = function(ns){
  return(seq(from=0,to=1,length.out = ns))
}

states = get.states(ns)

@@

Single Diffusion Steps
----------------------

Let's plot an arbitrary sender and receiver strategy, and apply some noise, step by step until convergence.

Parameters

@@

ns = 50
nm = 2
rounds = c(1,2,10,50,1000)
alphas = c(0.01,0.05,0.1,0.3)

plotSender = function(ns,nm,alphas,rounds){
  sen = rdirichlet(ns,rep(1,nm))
  senDF = data.frame(state = 0,
                     message = 0,
                     round = 0,
                     alpha = 0,
                     probability = 0)

  for (alpha in alphas){
    C = get.confusion.matrix(get.states(ns),alpha)
    newDF = data.frame(state = rep(get.states(ns),nm),
                       message = rep(1:2,each=ns),
                       round = rep(0,ns*nm),
                       alpha = rep(alpha,ns*nm),
                       probability = as.vector(sen))
    for (r in rounds){
      newC = power(C,r)
      newSen = get.nbr(sen,newC)
      newDF2 = data.frame(state = rep(get.states(ns),nm),
                          message = rep(1:2,each=ns),
                          round = rep(r,ns*nm),
                          alpha = rep(alpha,ns*nm),
                          probability = as.vector(newSen))
      newDF = rbind(newDF,newDF2)                   
      }
    senDF = rbind(senDF,newDF)
  }
  senDF = senDF[-1,]
  senDF$message = factor(senDF$message)
  senDF$alpha = factor(senDF$alpha)
  p = ggplot(senDF,aes(x = state, y = probability, color = message)) + 
      geom_point(size=0.5) + geom_line() + facet_grid(alpha~round) +
      scale_x_continuous(breaks=seq(0,1,1/2)) +
      theme(axis.text.x  = element_text(angle=90, vjust=0.5)) + 
      scale_y_continuous(breaks=seq(0,1,1/2))

  return(p)
}

plotReceiver = function(ns,nm,alphas,rounds){
  rec = rdirichlet(nm,rep(1,ns))
  recDF = data.frame(state = 0,
                     message = 0,
                     round = 0,
                     alpha = 0,
                     probability = 0)

  for (alpha in alphas){
    C = get.confusion.matrix(get.states(ns),alpha)
    newDF = data.frame(state = rep(get.states(ns),nm),
                       message = rep(1:2,each=ns),
                       round = rep(0,ns*nm),
                       alpha = rep(alpha,ns*nm),
                       probability = as.vector(t(rec)))
    for (r in rounds){
      newC = power(C,r)
      newRec = get.nbr(rec,newC)
      newDF2 = data.frame(state = rep(get.states(ns),nm),
                          message = rep(1:2,each=ns),
                          round = rep(r,ns*nm),
                          alpha = rep(alpha,ns*nm),
                          probability = as.vector(t(newRec)))
      newDF = rbind(newDF,newDF2)                   
      }
    recDF = rbind(recDF,newDF)
  }
  recDF = recDF[-1,]
  recDF$message = factor(recDF$message)
  recDF$alpha = factor(recDF$alpha)
  p = ggplot(recDF,aes(x = state, y = probability, color = message)) + 
      geom_point(size=0.5) + geom_line() + facet_grid(alpha~round) +
      scale_x_continuous(breaks=seq(0,1,1/2)) +
      theme(axis.text.x  = element_text(angle=90, vjust=0.5)) + 
      scale_y_continuous(breaks=seq(0,round(max(recDF$probability),2),round(max(recDF$probability),2)/2))
  return(p)
}

pSen = plotSender(ns,nm,alphas,rounds)
pRec = plotReceiver(ns,nm,alphas,rounds)
show(pSen)
show(pRec)

pdf(file = "paper/plots/confusion_sender.pdf", width = 7, height = 4)
show(pSen)
dev.off()

pdf(file = "paper/plots/confusion_receiver.pdf", width = 7, height = 4)
show(pRec)
dev.off()
